---
title: "Hotels"
author: "Bashayr Alghamdi"
date: "10/10/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 5 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
```

## Hotels booking analysis

this project about Hotel booking.
the dataset comes from an open hotel booking from [tidy tuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-02-11) demand dataset from Antonio,Almeida and Nunes, 2019.

```{r data set,}
hotels <- readr::read_csv("Data/hotels.csv")
```

(What are the reserarch questions that you can answer using this data set? What kind of insights can I gather using this data set? or predictions?)


### Data set

```{r}
glimpse(hotels)
```

Here the summary of the data set
```{r hotels summary}

summary(hotels) 

```

### missing value

When we focus on summary we will find the children column has 4 NA that means there is 4 missing values 

```{r}
summary(hotels$children)
```


```{r NA value,include=FALSE}
hotels_new <- hotels %>% 
  na.omit()
```

now we remove the NA value 

```{r}
summary(hotels_new$children)
```


## Date 

### Arrival date

we want gather the date of arrival in one column separated by "/"


```{r}

library(lubridate)


hotels <- hotels %>% 
  mutate(arrival_date = dmy(paste(
    arrival_date_day_of_month,
    arrival_date_month,
    arrival_date_year)))
hotels$arrival_date[10]

# hotels %>%
#   unite(arrival_data,
#         arrival_date_day_of_month,
#         arrival_date_month,
#         arrival_date_year,
#         sep="/")->hotels_date 

hotels_date$arrival_data[10]
#How can I change the month from name to number
```

### months 

Percentage of which month is there more reservation

```{r}
hotels %>% 
  group_by(arrival_date_month) %>% 
  summarise(count=n(),
            Percentage=(n()/nrow(hotels))*100)->months_count
months_count
```

```{r}
#Pie chart of parentage of all months
months_count
pie <- ggplot(data=months_count,aes(x="",y=Percentage,fill=arrival_date_month)) +
  geom_col(color= "white")+
  coord_polar("y",start = 0)+
  geom_text(aes(label=paste0(round(Percentage),"%")),
            position = position_stack(vjust=0.7))+
  theme_classic()+
  theme(axis.line= element_blank())
pie 
ggplot(data=months_count,aes(x=arrival_date_month,y=Percentage,fill=arrival_date_month))+
  geom_col(position = "dodge", shape = 16, alpha = 0.6)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

```

the result is August is the most month that had more reservation .in the other hand July is the lowest.


```{r}
hotels %>% 
  mutate(kids = children+babies) %>% 
  select(hotel,
         is_canceled,
         arrival_date_month,
         adults,
         kids,
         required_car_parking_spaces)->hotel_ds

```

her for each month what is the percentage between who has kids or not.

```{r}
hotel_ds %>% 
  filter(is_canceled == 0) %>% 
  mutate(have_kids=case_when(kids > 0 ~ "have kids",TRUE ~ "No kids"))->hotels_kids

hotels_kids %>%
  group_by(arrival_date_month,have_kids) %>% 
  summarise(count=n()) %>% 
  ggplot(aes(x = arrival_date_month, y = count, fill = have_kids)) +
  geom_bar(stat = "identity", shape = 16, alpha = 0.6)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") 

```

## car parking

How many customers that had kids take car parking space

```{r}
hotels %>% 
  select(adults, babies, children,required_car_parking_spaces) %>% 
  filter(babies+children != 0 & required_car_parking_spaces >= 1) %>% 
  summarise(count=n())#1,102

hotels %>% 
  select(adults, babies, children,required_car_parking_spaces) %>% 
  filter(babies+children != 0 & required_car_parking_spaces >= 1) %>% 
  mutate(number_of_customer=adults+babies+children) %>% 
  ggplot(aes(x = number_of_customer,y = required_car_parking_spaces))+
  geom_jitter(alpha = 0.25)+
  scale_y_continuous(limits = c(0,3),
                     expand = c(0,0)) +
  scale_x_continuous(limits = c(1,6),
                     expand = c(0,0))
```

How many customers take car parking space 

```{r}
hotels %>% 
  select(adults, babies, children,required_car_parking_spaces) %>% 
  filter(babies+children == 0 & required_car_parking_spaces >= 1) %>% 
  summarise(count=n())#6,314 

hotels %>% 
  select(adults, babies, children,required_car_parking_spaces) %>% 
  filter(babies+children == 0 & required_car_parking_spaces >= 1) %>% 
  mutate(number_of_customer=adults+babies+children) %>% 
  ggplot(aes(x = number_of_customer,y = required_car_parking_spaces))+
  geom_jitter(alpha = 0.25)+
  scale_y_continuous(limits = c(0,10),
                     expand = c(0,0)) +
  scale_x_continuous(limits = c(0,5),
                     expand = c(0,0))
```

## Night

what the total of the night that the customer stayed 

```{r}
#calculate total of the night that the customer stayed  
hotels %>% 
  select(stays_in_weekend_nights,stays_in_week_nights) %>% 
  mutate(stays_nights = stays_in_weekend_nights + stays_in_week_nights) -> hotels_nights
hotels_nights
```







