---
title: "Hotels"
author: "Bashayr Alghamdi"
date: "10/10/2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

## Hotels booking analysis

this project about Hotel booking.
the dataset comes from an open hotel booking from [tidy tuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-02-11) demand dataset from Antonio,Almeida and Nunes, 2019.

```{r data set}
# Read in the data (csv format):
hotels <- readr::read_csv("Data/hotels.csv")
```


### Data set

```{r}
#describe each column what values it has
glimpse(hotels)
```
In this research we will answer some questions about customer reservation for each month.

So we do not need to use all of these columns, we need some specific columns


```{r}
#select the columns that we need
hotels %>% 
  select(hotel,is_canceled,arrival_date_month,adults,children,babies,customer_type,stays_in_weekend_nights,stays_in_week_nights)->hotels_ss #create new subset of data set 



```

## Descriptive statice


Here the summary of the data set
```{r hotels summary}

summary(hotels_ss) 

```

### missing value

When we focus on summary we will find the children column has 4 NA that means there is 4 missing values 

```{r}
summary(hotels_ss$children)
```


```{r NA value,include=FALSE}
hotels_ss <- hotels_ss %>% 
  na.omit()
```

now we remove the NA value 

```{r}
summary(hotels_ss$children)
```


### data visiuaiztion 


Percentage of which month is there more reservation

```{r}
hotels_ss %>% 
  group_by(arrival_date_month) %>% 
  summarise(count=n(),
            Percentage=(n()/nrow(hotels))*100)->months_count

```

```{r}
#Pie chart of parentage of all months

ggplot(data=months_count,aes(x=arrival_date_month,y=Percentage,fill=arrival_date_month))+
  geom_col(position = "dodge", shape = 16, alpha = 0.6)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")


```

the result is August is the most month that had more reservation .in the other hand July is the lowest.


```{r}
hotels %>% 
  mutate(kids = children+babies) %>% 
  select(hotel,
         is_canceled,
         arrival_date_month,
         adults,
         kids,
         required_car_parking_spaces)->hotel_ds

```

her for each month what is the percentage between who has kids or not.

```{r}
hotel_ds %>% 
  filter(is_canceled == 0) %>% 
  mutate(have_kids=case_when(kids > 0 ~ "have kids",TRUE ~ "No kids"))->hotels_kids

hotels_kids %>%
  group_by(arrival_date_month,have_kids) %>% 
  summarise(count=n()) %>% 
  ggplot(aes(x = arrival_date_month, y = count, fill = have_kids)) +
  geom_bar(stat = "identity", shape = 16, alpha = 0.6)+
  geom_text(aes(label = count), size = 3, hjust = 0.5, vjust = 3, position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") 

```











